
% Wed 18 Mar 2015
% section

Intermolecular interactions have an important effect on the physical and chemical properties of condensed chemical systems, especially where non-covalent interactions dominate. 
The calculation of the interaction energy requires accounting for the Coulomb interaction that dictates much of the intermolecular dynamics of ions, as well as other non-covalent interactions such as hydrogen-bonding, $\pi$-$\pi$ stacking, van der Waals forces, etc.
\cite{Wendler2012a, Bedrov2010a, Izgorodina2011a}
The noncovalent interactions in ionic liquids are often dominated by electrostatics (Coulomb), dispersion and induction (also known as polarization), as well as exchange-repulsion to a smaller extent. 
\cite{Izgorodina2011b, Izgorodina2014a}
The complex interplay of all these interactions means that characterising the intermolecular dynamics of ionic liquids is a challenging task. 
\cite{Izgorodina2011a}


% why is it state of the art, reproduce CC energies
% Robert Moszynski?? Tatiana Korona, both book sections
% read Parker2014
% Korona1997 (helium),  Sinnokrot2004a, Sinnokrot2006a (benzene dimers, pi-pi interactions compared with CCSD(T))

% no BSSE
Symmetry-adapted perturbation theory (SAPT)
\cite{Misquitta2008a, Misquitta2008b, Parker2014a}
is the state-of-the-art method for calculating the strength of intermolecular interactions.
SAPT decomposes the total interaction energy into the five fundamental components: electrostatics, exchange, induction, dispersion and charge-transfer.
%Out of these components, charge-transfer is accounted for as part of the induction energy within the formulation.
The analysis of these components provides important insight into how they affect the structure and physicochemical properties of the chemical system in consideration.
\cite{Stone1996a, Turney2012a}
It has been used to study components of the interaction energy, such as the dispersion,
\cite{Misquitta2005a} 
induction for small organic molecules and organic crystals,
\cite{Misquitta2008a, Misquitta2008b, Welch2008a}
charge-transfer,
\cite{Karthikeyan2011a}
and hydrogen-bonding,
\cite{Milet1999a}
as well as the well-studied helium dimer potential
\cite{Korona1997a}
and $\pi \text{--} \pi$ interactions in benzene.
\cite{Sinnokrot2004a, Sinnokrot2006a}
However, while accurate, this method is very expensive computationally, scaling to at least $N^5$ when truncated at second order.
Electrostatics and exchange occur in the first order of the interaction potential. 
Second order SAPT has the dispersion and induction terms.


The general effective fragment potential (EFP2, referred to as simply EFP in the rest of the text) method was developed by Gordon et al. 
\cite{Jensen1998a, Gordon2001a, Gordon2009a, Mullin2009a, Gordon2012a} 
as a computationally inexpensive method to model energetic components of intermolecular interactions. 
This method was originally created to model solvents (water),
\cite{Day1996a, Chen1996a, Adamovic2006a} 
but has been later generalised for non-covalently bound complexes.
\cite{Gordon2007a, Ghosh2010a}
It belongs to class of fragmentation methods, and is an \emph{ab-initio} based method, without any empirical parameters, with each term being developed independently of the rest. 
Each term in the EFP method represents an individual fundamental component of interaction energy such as electrostatics, exchange, polarization, dispersion and charge transfer. 
Thus it calculates the interaction energy as a sum of terms directly comparable with correlated wavefunction-based methods such as coupled cluster and perturbation theory.
Some applications of the EFP method in the literature include solvent systems
\cite{Chen1996a, Day1996a, Day1997a, Krauss1997a, Merrill1998a, Day2000a, Adamovic2006a}
solvent-induced shifts in the electronic spectra of uracil,
\cite{DeFusco2011a}
spectroscopy of enzyme active sites,
\cite{Krauss1995a, Krauss1998a}
DNA,
\cite{Ghosh2010a}
hydrogen bonding,
\cite{Jensen1994a}
and other non-covalently bound systems
\cite{Gordon2009a, Gordon2013a, Wladkowski1995a}
% specific about comparison with CCSD and MP2, cite papers and give examples of systems compared

EFP has been applied to reliably study a broad range of intermolecular complexes. For example, EFP relative energies of hydrogen bonded complexes in $ (\text{MeOH/H}_2\text{O})_{n} $ clusters, for $ n = 2, \ldots, 8 $ were in good agreement with MP2.
\cite{Adamovic2006a}

In the case of a number of configurations of styrene dimers, exhibiting hydrogen bonding and  $\pi$-$\pi$ stacking interactions, EFP generally reproduced MP2 geometries and binding energies with the exception of one system.
\cite{Adamovic2006b}


% Slipchenko 
In the work of Slipchenko et al., the EFP method was found to perform very well for the $\pi$-$\pi$ stacked benzene dimers in three different configurations.
For the most challenging configuration, the parallel-displaced benzene dimer, the difference between EFP and CCSD(T)/aug-cc-pVQZ was only 0.3 kcal/mol, which is a smaller error than typical MP2 at a much lower computational cost.
In addition, the EFP method was shown to recover around 70\% of the charge penetration energy, arising due to the decrease of the classical electrostatic energy as a result of electron density overap.
\cite{Slipchenko2007a}


% benzene-pyridine interactions, Smith QA
In further work examining $\pi$-$\pi$ stacking
\cite{Smith2008a}, 
studying dimers of benzene, pyridine and benzene-pyridine, the EFP method was found to be in good agreement with CCSD(T), MP2 and SAPT2 results for the calculation of potential energy surfaces and interaction energy components.
The average root-mean-square deviation between EFP and CCSD(T) was 0.49 kcal/mol with a range of 0.31 kcal/mol to 0.66 kcal/mol.
\cite{Smith2011a,Slipchenko2009a}


% DNA bases, Smith QA Gordon Slipchenko
For hydrogen-bonded and stacked DNA base pairs, Smith et al. found fair to excellent agreement between EFP and CCSD(T) results, and further emphasised the importance of a method that accurately treats the dispersion and polarisation terms.
While the electrostic term dominates in hygdrogen-bonded systems, its contribution to the binding energy is largely offset by exchange-repulsion, meaning dispersion and polarisation interactions play an important part in the stabilisation of these systems.
EFP is able to account for these interactions at a much lower cost compared to MP2, CCSD(T) and SAPT, while being in reasonable agreement with these methods.
\cite{Smith2011b}


% Hands Slipchenko tert-butanol water mixtures
Hands and Slipchenko have also pioneered a study validating the EFP method for polar heterogenic systems,
\cite{Hands2012a}
using this method to investigate \emph{tert}-butanol water mixtures with particular attention to hydrogen bonding.
While the EFP method has been shown to compare well with other higher level \emph{ab-initio} methods for small clusters, much of its potential lies in the possibility of applying it to large clusters, which are typically treated with force fields and molecular dynamics at the time of writing.
They found that while EFP accurately reproduced the experimental structure of pure water and water--\emph{tert}-butanol mixtures, hydrogen-bonding patterns are very sensitive to the model potential used.
Furthermore, the significance of the polarisation energy was emphasised in the bulk compared to dimer systems where electrostatics and repulsion terms dominate.


Flick et al. undertook a systematic study on the performance of EFP for the prediction of interaction energies in non-covalently bound complexes
\cite{Flick2012a}.
The results were compared against a raft of semi-empirical and correlated wavefunction-based methods as well as methods of density functional theory.
In their study, the two well-known databases, S22 and S66 of Hobza et al.
\cite{Jurecka2006a, Rezac2011a}    % get these on Mon
were used.
These benchmark databases contain a well-balanced series of intermolecular complexes such as complexes between small molecules such as water and ammonia, DNA base pairs, and amino acid pairs.


The EFP method reproduced the interaction energies rather well, with a mean unsigned error of 2.5 \enUnit~ for the S66 set and 3.8 \enUnit~ for the S22 set.
This corresponds to a relative error of 11-12\% in interaction energy.
For the S22 set, they found that the strongly interacting hydrogen bonding complexes had below average accuracy, with an underestimation of 10-15\% in the electrostatic energy, and attributed to incomplete treatment of charge-penetration.
The electrostatic term for mixed complexes was better treated since the magnitude of the interaction was smaller, and charge-penetration effects did not dominate.
This is not observed in the S66 set as the hydrogen bonding complexes in S66 have weaker interactions compared to the ones in S22.
Ionic liquids in general have strong non-covalent interactions, and many ILs have significant hydrogen bonding.
In light of this, EFP might underestimate the strong interactions present in ILs.


Induction (polarisation) was another term that was significantly underestimated (4-5 kcal/mol) for hydrogen-bonded complexes.
The relative energies for dispersion-dominated complexes have even larger errors, but due to the small magnitudes of their polarisation energies the associated absolute errors are not large (largest ~2 kcal/mol).
This is attributed to the limitations of the multipole approximation for dispersion-dominated systems.
Furthermore, the charge-transfer energy, which is included in the SAPT induction term, is neglected when EFP polarisation is compared with SAPT induction.
Charge-transfer is negligible for dispersion dominated neutral complexes, but may play a more important role in strongly hydrogen-bonded complexes.
In ionic liquids, charge-transfer has been demonstrated to occur regularly and is seen to be an important contribution,
\cite{Izgorodina2011a}
but the magnitude of its stabilisation effects on the total interaction energy have not been verified.

The underestimation of the electrostatic and polarisation terms discussed previously is offset in part by underestimation in the exchange-repulsion term, with typical relative errors of 10-20\%.
Once again, the largest errors occur for the complexes where hydrogen-bonding dominates.
This underestimation is hypothesised to arise from the neglect of correlation effects in EFP.


On the other hand, Flick et al. showed that EFP is quite good at giving a balanced description for different types of interactions, performing well with dispersion-dominated complexes and complexes with a mix of interaction types.
% maybe some more about Flick's results
The accuracy of EFP approached that of second order M\"{o}ller--Plesset perturbation theory method, MP2, which is widely used for studying intermolecular complexes.
They found that the main sources of error in EFP come from the underestimation of the polarisation and Coulomb components, which they attributed to insufficient treatement of short-range charge-penetration effects.
This is partially offset by the underestimation in the exchange-repulsion term.

They also used EFP to predict the intermolecular separation between the complexes, while keeping the EFP fragments rigid.
Upon comparison with the ab-initio geometry optimisations from CCSD(T) calculations, they found that in general EFP tends to overestimate the intermolecular distance, suggesting an underestimation of the attractive forces or an overestimation of repulsion.



EFP is an attractive option for studying ionic liquids since it decomposes the interaction energy into physically meaningful components.
Energy decomposition provides further insights on the relationship between calculated energies and physico-chemical properties.
Understanding this relationship will tremendously aid in the prediction of properties and design of new ionic liquids.
Furthermore, EFP is computationally cost-effective, scaling as $N^2$.
\cite{Flick2012a}


However, to date, no systematic study has been done on the suitability of the EFP method for charged species like ionic liquids. 
While this method was not originally designed for charged species, its computational efficiency renders it a good candidate. 
Furthermore, it's decomposition of the interaction energy into its fundamental components is a valuable feature for the study of the relationship between energetics and physical properties. 
Based on the findings of Flick et al., the three terms, Coulomb, exchange-repulsion and polarisation, are expected to show larger errors for ionic liquids. 
In addition charge-transfer might also be not treated accurately due to significant orbital overlap in ionic liquids, as is indicated by fractional charges that IL ions adopt.
\cite{Schmidt2010a, Dommert2012a, Dommert2014a, Rigby2013a, Wendler2012a}
The charge-transfer term has been shrouded in controversy, and it is still unclear whether it should be included as a stabilising energy term or whether an actual transfer of electronic charge is necessary for performing calculations involving accurate structure and dynamics.
\cite{Ramesh2008a, Robertson2002a, Thompson2000a, Lee2011a, Piquemal2006a, Kumar2011a}
This is due to charge-transfer being inherently tied with polarisation, leading to varied formulations ranging from charge-transfer being an artifact of the incompleteness of the basis set, to theories such as natural energy decomposition analysis,
\cite{Schenter1996a, Glendening2005a}
where it is a prevailing force in intermolecular attraction.
Accurate treatment of these components is crucial in ionic liquids, where these are expected to constitute the bulk of the total interaction energy.
This study identifies how well EFP performs for ionic liquids for the prediction of interaction energies in single ion pairs of ionic liquids.
The test set consists of a large number of cation-anion combinations at various configurations.
Imidazolium and pyrrolidinium based cations were coupled with routinely used anions, such as chloride and tetrafluoroborate.


A brief overview of the theory is given in the next section. 
After that the computational methodology is outlined in Section \ref{sec:method}, followed by discussion of the results in Section \ref{sec:results}.
% literature search on SAPT and EFP for studying intermolecular complexes
