
\documentclass[journal=jctcce]{achemso}

\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{color}


% for esi.tex
\usepackage{pdflscape}
\usepackage{multirow}
% too many unprocessed floats error
\usepackage{morefloats}
\errorcontextlines 10000

%\usepackage{floatrow}
%\floatsetup[table]{font=footnotesize}
%\floatsetup[table]{capposition=top}

%% too many unprocessed floats error
\usepackage{morefloats}
\errorcontextlines 10000


\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{appendix}

\usepackage{xstring}

% for symbols in footnotes
\usepackage[symbol*]{footmisc}

\usepackage{natbib}

\newcommand{\enUnit}{kJ$ \cdot \text{mol}^{-1}$}

% commands for all the anions
\newcommand{\bfl}{ \ensuremath{ \text{BF}_4^- } }
\newcommand{\br}{ \ensuremath{ \text{Br}^- } }
\newcommand{\cl}{ \ensuremath{ \text{Cl}^- } }
\newcommand{\dca}{ \ensuremath{ \text{Dca}^- } }
\newcommand{\mes}{ \ensuremath{ \text{Mes}^- } }
\newcommand{\ntf}{ \ensuremath{ \text{NTf}_{2}^{-} } }
\newcommand{\pf}{ \ensuremath{ \text{PF}_{6}^{-} } }
\newcommand{\tos}{ \ensuremath{ \text{Tos}^- } }

\newcommand{\bflb}{ $ [ \text{BF}_4 ] $ }
\newcommand{\brb}{ $ [ \text{Br} ] $ }
\newcommand{\clb}{ $ [ \text{Cl} ] $ }
\newcommand{\dcab}{ $ [ \text{Dca} ] $ }
\newcommand{\mesb}{ $ [ \text{Mes} ] $ }
\newcommand{\ntfb}{ $ [ \text{NTf}_{2} ] $ }
\newcommand{\pfb}{ $ [ \text{PF}_{6} ] $ }
\newcommand{\tosb}{ $ [ \text{Tos} ] $ }

% for cations
\newcommand{\cat}[2]{ $ \text{C}_{#2}\text{#1}^{+} $ }
\newcommand{\catb}[2]{ $ [ \text{C}_{#2}\text{#1} ] $ }

% for ion pairs
\newcommand{\ipair}[3]{
    \IfEqCase{#3} {
        {bfl} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{BF}_4 ]}}  
        {br} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{Br} ]}}
        {cl} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{Cl} ]}}
        {dca} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{Dca} ]}}
        {mes} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{Mes} ]}}
        {ntf} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{NTf}_{2} ]}} 
        {pf} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{PF}_6 ]}}
        {tos} {\ensuremath{[ \text{C}_{\text{#2}}\text{#1}] [ \text{Tos} ]}}   
    } 
    [ \PackageError{ipair}{Undefined option (anion) to ipair: #3}{} ]
}

% for energies
% 2nd arg optional; \energ[Ind]{EFP}
\newcommand{\energ}[2][]{ \ensuremath{ E^{\text{#1}}_{\text{#2}} }}

\title{Comparison of the Effective Fragment Potential Method with Symmetry-Adapted Perturbation Theory in the Calculation of Intermolecular Energies for Ionic Liquids}
%\date{}
\author{Samuel Y. S. Tan}
\email{samuel.tan@monash.edu} 
\author{Ekaterina I. Izgorodina}
\email{katya.pas@monash.edu}
\affiliation[Monash University]
    {School of Chemistry, Monash University, 17 Rainforest Walk, Clayton, Victoria 3800, AUSTRALIA}

\begin{document}

\maketitle

\begin{abstract}
    The effective fragment potential (EFP) method that decomposes the interaction energy as a sum of the five fundamental forces---electrostatic, exchange-repulsion, polarisation, dispersion and charge-transfer---was applied to a large test set of ionic liquid ion pairs and compared against the state-of-the-art method, Symmetry-Adapted Perturbation Theory (SAPT).
    The ion pairs include imidazolium and pyrrolidinium cations combined with anions that are routinely used in the field of ionic liquids.
    The aug-cc-pVDZ, aug-cc-pVTZ, and 6-311++G(d,p) basis sets were used for EFP, while SAPT2+3/aug-cc-pVDZ provided the benchmark energies.
    Differences between the two methods were found to be large, and strongly dependent on the anion type. 
    For the aug-cc-pVTZ basis set, which produced the least errors, average relative errors were between 2.3--18.4\% for pyrrolidinium ion pairs and 2.1--27.7\% for imidazolium ion pairs for each individual energetic component (excluding charge-transfer) as well as the total interaction energy.
    Charge transfer gave the largest relative errors, 56\% and 63\% on average for pyrrolidinium- and imidazolium-based ion pairs, respectively.
    Scaling of the EFP components against SAPT2+3 showed improvement for polarisation (induction) and dispersion terms, thus indicating potential for the development of cost-effective alternatives for intermolecular induction and dispersion potentials for ionic liquids. 
    
    
\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{./tex/intro.tex}

\section{Theoretical Background}
\label{sec:bkgrd}
\input{./tex/bkgrd.tex}

\section{Methodology}
\label{sec:method}
%%Note that these numbers do not include \ipair{mpyr}{3}{dca} (p5) which is an outlier (0.331 \enUnit).
\input{./tex/method.tex}

\section{Results and Discussion}
\label{sec:results}

% due to MAE SD formulas etc
\input{./tex/ccsd.tex}

\subsection{Differences between SAPT2+3 and EFP}
\label{subsec:comparison}
% absolute and relative (ratio diffferences)
\input{./tex/comparison.tex}

\subsection{Correlation between SAPT and EFP}
\label{subsec:corr-elst}
%% correlations and scaling results, both indiv and ratio-wise
\input{./tex/corr_elst.tex}

%\subsection{Charge-transfer}
\input{./tex/geodesic.tex}

\section{Conclusion}
\label{sec:conclusion}
\input{./tex/conclusion.tex}

%\section*{Acknowledgements}
\begin{acknowledgement}
The authors gratefully acknowledge a generous allocation of computer resources through the Monash \emph{e}Research
Centre and the National Computational Infrastructure. This
work is generously supported by the Australian Research Council through a Discovery Project Grant and a Future Fellowship for EII.
\end{acknowledgement}

\begin{suppinfo}
    The supplementary information includes data not displayed in the main text.
    The statistics on the differences between the two SAPT2+3 basis sets, aug-cc-pVDZ and aug-cc-pVTZ, linear regression statistics for the other two EFP basis sets, aug-cc-pVDZ and 6-311++G(d,p), as well as the raw energies from both SAPT2+3 and EFP for each ion pair are all presented here.
\end{suppinfo}

%\input{./tex/esi.tex}

%\renewcommand{\bibfont}{\footnotesize}
%\printbibliography

% for natbib
\clearpage
\bibliography{sapt_efp}

\end{document}
